---
title: "Model Comparison Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    self_contained: true
  word_document:
    toc: true
    toc_depth: 3
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
params:
  raw_data:        NULL
  roles:           NULL
  col_types:       NULL
  selected_models: NULL
  anova_wide:      NULL
  rmse_df:         NULL
  vif_df:          NULL
  lsmeans_list:    NULL
  effects_list:    NULL
  mc_results_list: NULL
  mc_on:           FALSE
  mc_terms:        NULL
  mc_methods:      NULL
  repro_code:      NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 7,
  fig.height = 4.5,
  dpi = 150
)

library(ggplot2)
library(dplyr)
library(flextable)

# Convenience: safe flextable from data.frame
ft <- function(df, caption = NULL) {
  if (is.null(df) || nrow(df) == 0) {
    return(flextable::flextable(data.frame(Note = "No data available.")))
  }
  tbl <- flextable::flextable(as.data.frame(df))
  tbl <- flextable::autofit(tbl)
  tbl <- flextable::theme_vanilla(tbl)
  if (!is.null(caption)) {
    tbl <- flextable::set_caption(tbl, caption)
  }
  tbl
}

model_colours <- function(n) {
  palette <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728",
               "#9467bd", "#8c564b", "#e377c2", "#7f7f7f",
               "#bcbd22", "#17becf")
  rep_len(palette, n)
}
```

# 1. Raw Data Preview

The first rows of the uploaded dataset are shown below.

```{r raw-data}
if (!is.null(params$raw_data) && nrow(params$raw_data) > 0) {
  ft(head(params$raw_data, 20), caption = "Raw data (first 20 rows)")
} else {
  cat("No data available.")
}
```

# 2. Column Role and Type Assignments

```{r roles}
if (!is.null(params$roles) && length(params$roles) > 0) {
  role_df <- data.frame(
    Column = names(params$roles),
    Role   = unlist(params$roles),
    stringsAsFactors = FALSE
  )
  if (!is.null(params$col_types) && length(params$col_types) > 0) {
    type_df <- data.frame(
      Column = names(params$col_types),
      Type   = unlist(params$col_types),
      stringsAsFactors = FALSE
    )
    role_df <- merge(role_df, type_df, by = "Column", all.x = TRUE)
    role_df$Type[is.na(role_df$Type)] <- "\u2014"
  }
  ft(role_df, caption = "Column role and type assignments")
} else {
  cat("No role assignments available.")
}
```

# 3. ANOVA Table (Type III)

Models included in this report: `r paste(params$selected_models, collapse = ", ")`.

```{r anova-table}
if (!is.null(params$anova_wide) && nrow(params$anova_wide) > 0) {
  num_cols <- sapply(params$anova_wide, is.numeric)
  anova_fmt <- params$anova_wide
  anova_fmt[num_cols] <- lapply(anova_fmt[num_cols], function(x) round(x, 4))
  ft(anova_fmt, caption = "ANOVA Type III results (F values and p-values per model)")
} else {
  cat("No ANOVA results available.")
}
```

*Note: p-values < 0.05 indicate statistically significant terms. Rows ordered: factors, blocks, covariates, residuals.*

## 3.1 Model Fit Summary (RMSE / R²)

```{r rmse-table}
if (!is.null(params$rmse_df) && nrow(params$rmse_df) > 0) {
  ft(params$rmse_df, caption = "Model fit statistics")
} else {
  cat("No model fit statistics available.")
}
```

## 3.2 Collinearity (VIF)

```{r vif-table}
if (!is.null(params$vif_df) && nrow(params$vif_df) > 0) {
  vif_tbl <- ft(params$vif_df, caption = "VIF (cross-model comparison)")
  model_cols <- setdiff(names(params$vif_df), "Term")
  for (mc in model_cols) {
    vals <- params$vif_df[[mc]]
    high_i   <- which(!is.na(vals) & vals > sqrt(5) & vals <= sqrt(10))
    vhigh_i  <- which(!is.na(vals) & vals > sqrt(10))
    if (length(high_i) > 0)
      vif_tbl <- flextable::bg(vif_tbl, i = high_i, j = mc, bg = "#fff3cd")
    if (length(vhigh_i) > 0)
      vif_tbl <- flextable::bg(vif_tbl, i = vhigh_i, j = mc, bg = "#f8d7da")
  }
  vif_tbl
} else {
  cat("No VIF results available (models may have fewer than 2 predictors).")
}
```

# 4. Effects Plots

Factor terms show LS means (\u00b1 SE). Covariates show partial effects with 95% confidence bands. Both include adjusted data overlays where other model terms have been removed.

```{r effects-factor-plots, results='asis', fig.width=7, fig.height=4}
if (!is.null(params$lsmeans_list) && length(params$lsmeans_list) > 0) {
  sec_idx <- 0L
  for (spec in names(params$lsmeans_list)) {
    ls_df <- params$lsmeans_list[[spec]]
    if (is.null(ls_df) || nrow(ls_df) == 0) next

    sec_idx <- sec_idx + 1L
    cat(paste0("\n\n## 4.", sec_idx, " LS Means: ", spec, "\n\n"))

    spec_terms <- strsplit(spec, ":")[[1]]
    x_col <- spec_terms[1]
    y_col <- if ("emmean" %in% names(ls_df)) "emmean" else names(ls_df)[2]
    se_col <- if ("SE" %in% names(ls_df)) "SE" else NULL

    colours <- model_colours(length(unique(ls_df$model)))

    p <- ggplot(ls_df, aes_string(x = x_col, y = y_col,
                                   colour = "model", group = "model")) +
      geom_point(size = 3, position = position_dodge(0.3))

    if (!is.null(se_col)) {
      p <- p + geom_errorbar(
        aes_string(ymin = paste0(y_col, " - ", se_col),
                   ymax = paste0(y_col, " + ", se_col)),
        width = 0.2, position = position_dodge(0.3)
      )
    }

    p <- p +
      scale_colour_manual(values = colours) +
      labs(title = paste("LS Means \u2014", spec),
           x = x_col, y = "LS Mean", colour = "Model") +
      theme_minimal(base_size = 11)

    print(p)
    cat("\n\n")

    # LS means table
    cat("\n**LS Means values:**\n\n")
    num_cols <- sapply(ls_df, is.numeric)
    ls_fmt <- ls_df
    ls_fmt[num_cols] <- lapply(ls_fmt[num_cols], function(x) round(x, 4))
    print(ft(ls_fmt))
    cat("\n\n")
  }
} else {
  cat("No LS means results available.")
}
```

```{r effects-cov-plots, results='asis', fig.width=7, fig.height=4}
if (!is.null(params$effects_list) && length(params$effects_list) > 0) {
  sec_base <- if (!is.null(params$lsmeans_list)) length(params$lsmeans_list) else 0L
  for (i in seq_along(params$effects_list)) {
    cov <- names(params$effects_list)[i]
    eff_df <- params$effects_list[[cov]]
    if (is.null(eff_df) || nrow(eff_df) == 0) next

    cat(paste0("\n\n## 4.", sec_base + i,
               " Partial Effect: ", cov, "\n\n"))

    colours <- model_colours(length(unique(eff_df$model)))

    p <- ggplot(eff_df, aes(x = x, y = fit, colour = model, fill = model)) +
      geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.15, colour = NA) +
      geom_line(linewidth = 1) +
      scale_colour_manual(values = colours) +
      scale_fill_manual(values = colours) +
      labs(title = paste("Partial effect of", cov),
           subtitle = "Other covariates at mean; factors at reference level",
           x = cov, y = "Predicted response", colour = "Model") +
      guides(fill = "none") +
      theme_minimal(base_size = 11)

    print(p)
    cat("\n\n")
  }
} else {
  cat("No covariate effects data available.")
}
```

# 5. Multiple Comparisons

```{r mc-section, results='asis', fig.width=7, fig.height=4}
if (isTRUE(params$mc_on) &&
    !is.null(params$mc_results_list) &&
    length(params$mc_results_list) > 0) {

  mc_keys <- names(params$mc_results_list)
  sec_idx <- 0L

  for (key in mc_keys) {
    mc_df <- params$mc_results_list[[key]]
    if (is.null(mc_df) || nrow(mc_df) == 0) next

    sec_idx <- sec_idx + 1L
    parts  <- strsplit(key, "__")[[1]]
    mname  <- parts[1]
    spec   <- if (length(parts) >= 2) parts[2] else "unknown"
    method <- if (length(parts) >= 3) parts[3] else "unknown"

    cat(paste0("\n\n## 5.", sec_idx, " Comparisons: ", spec, " \u2014 ", method,
               " (", mname, ")\n\n"))

    est_col <- if ("estimate"   %in% names(mc_df)) "estimate"   else names(mc_df)[2]
    se_col  <- if ("SE"         %in% names(mc_df)) "SE"         else
               if ("std.error"  %in% names(mc_df)) "std.error"  else NULL

    if (!"contrast" %in% names(mc_df))
      mc_df$contrast <- as.character(seq_len(nrow(mc_df)))

    p <- ggplot(mc_df, aes_string(x = "contrast", y = est_col)) +
      geom_point(size = 2.5, colour = "#1f77b4") +
      geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50")

    if (!is.null(se_col)) {
      p <- p + geom_errorbar(
        aes_string(ymin = paste0(est_col, " - 1.96*", se_col),
                   ymax = paste0(est_col, " + 1.96*", se_col)),
        width = 0.3, colour = "#1f77b4"
      )
    }

    p <- p +
      coord_flip() +
      labs(title = paste("Pairwise estimates \u2014", spec, "(", method, ")"),
           x = "Contrast", y = "Estimate \u00b1 95% CI") +
      theme_minimal(base_size = 11)

    print(p)
    cat("\n\n")

    num_cols <- sapply(mc_df, is.numeric)
    mc_fmt <- mc_df
    mc_fmt[num_cols] <- lapply(mc_fmt[num_cols], function(x) round(x, 4))
    print(ft(mc_fmt, caption = paste("Comparison results \u2014", spec, "/", method, "/", mname)))
    cat("\n\n")
  }
} else {
  cat("Multiple comparisons were not enabled or no results available.")
}
```

# 6. Residual Diagnostics

*Residual diagnostic plots are generated interactively in the app (Residuals tab).
The models included in this report are listed below for reference.*

```{r residuals-note}
if (length(params$selected_models) > 0) {
  note_df <- data.frame(
    `#`    = seq_along(params$selected_models),
    Model  = params$selected_models,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  ft(note_df, caption = "Models in report")
} else {
  cat("No models available.")
}
```

# 7. Reproducible Code {.tabset}

This section contains R code to reproduce the analysis outputs in this report. Copy and paste into an R session with the DoEWorkbench package installed.

```{r repro-code, results='asis', echo=FALSE}
if (!is.null(params$repro_code) && nchar(params$repro_code) > 0) {
  cat("\n```r\n")
  cat(params$repro_code)
  cat("\n```\n")
} else {
  cat("*Reproducible code not available for this report format.*")
}
```

---

*Report generated by DoE Workbench on `r format(Sys.time(), "%d %B %Y at %H:%M")`.*
